---
title: "Homework - 4.2"
author: "By 刘慧航 SA18017026"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bootstrap)
library(emplik)
library(plotrix)
data(law)
```

## Problem

使用bootstrap包里的law数据, 使用经验似然方法绘制LSAT和GPA的均值的50%,90%,95%和99%置信区域, 并与正态分布下的置信区域相对比. （利用Owen主页上的EL程序）

## Solution

I use **el.test** function in  **emplik** package to calculate empirical likelihood ratio. Code is imitated from Stackflow as following.

And code of confidence regions under normal assumption is calculate imitated from Weiyu Li.

## Code 

```{r hw4}
# data preparation
num_grids <- 100
x <- law$LSAT
y <- law$GPA
df <- cbind(x,y)
myresult <- matrix(NA, ncol = num_grids, nrow = num_grids)

# get limit of data
mu_x <- seq(min(x), max(x), length.out = num_grids)
mu_y <- seq(min(y), max(y), length.out = num_grids)

for(i in 1:num_grids)
  for(j in 1:num_grids) 
    # calculate log likelihood ratio by el.test function in emplik package
    # myresult[i,j] <- -0.5 * el.test(x= df, mu = c(mu_x[i], mu_y[j]))$"-2LLR"
    myresult[i,j] <- el.test(x= df, mu = c(mu_x[i], mu_y[j]))$"-2LLR"

# draw means of data
plot(mean(x), mean(y), cex = 2, col = "brown", pch = 19, xlab = "LSAT", ylab = "GPA", main = "Empirical Likelihood Contours", xlim=c(560,640), ylim=c(2.8,3.4))

# Query Graphical
par(new=TRUE)   

# draw contour of confidence regions using empirical likelihood by emplik package, in red
# contour(mu_x,mu_y, exp(-0.5*myresult), levels=c(0.5,0.1,0.05,0.01), col="black", add=TRUE)
contour(mu_x,mu_y, myresult, levels=c(qchisq(0.5,2),qchisq(0.9,2),qchisq(0.95,2),qchisq(0.99,2)), col="red", add=TRUE)

# confidence regions under normal assumption in blue.
# draw by plotrix package
# ref: liweiyu
n <- nrow(law)
p <- 2
sig <- cov(law)
values <- eigen(sig)$values
vectors <- eigen(sig)$vectors
c<-(n-1)*p/(n-p)*qf(0.5,p,n-p)
draw.ellipse(mean(x),mean(y),a=sqrt(values[1]*c/n), b=sqrt(values[2]*c/n),angle=atan(vectors[2]/vectors[1]),deg=FALSE,border='blue')
c<-(n-1)*p/(n-p)*qf(0.9,p,n-p)
draw.ellipse(mean(x),mean(y),a=sqrt(values[1]*c/n), b=sqrt(values[2]*c/n),angle=atan(vectors[2]/vectors[1]),deg=FALSE,border='blue')
c<-(n-1)*p/(n-p)*qf(0.95,p,n-p)
draw.ellipse(mean(x),mean(y),a=sqrt(values[1]*c/n), b=sqrt(values[2]*c/n),angle=atan(vectors[2]/vectors[1]),deg=FALSE,border='blue')
c<-(n-1)*p/(n-p)*qf(0.99,p,n-p)
draw.ellipse(mean(x),mean(y),a=sqrt(values[1]*c/n), b=sqrt(values[2]*c/n),angle=atan(vectors[2]/vectors[1]),deg=FALSE,border='blue')
```

To draw the confidence region by EL. 
I firstly generate a grid with 100 bins according to the boundary of data. 
Then calculate $-2 log\mathcal{R(\theta)}$ at each data point $\theta$ within the grid.
After that, we can draw contour at different levels which is calculated by quantiles of $\chi^2_{(2)}$.

To draw the confidence ellipse under normal assumption. 
I imitate the code by Liweiyu. The radii and the angle of the confidence ellypses is calculated by eigen decomposition.
