---
title: "Homeword 11"
author: "Liu Huihang"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1

For the data set *faithful* in R, test whether the *eruptions* and *waiting* are independent of each other. 

## Answer 1

Assume that *eruptions*$=X\sim f_1(x)$ and  *waiting*$=Y\sim f_2(y)$.
We are going to do the following test
\begin{equation}
  H_0: \; f(x,y) = f_1(x) f_2(y) \Leftrightarrow H_1: \; f(x,y) \neq f_1(x) f_2(y)
\end{equation}
and observe that
\begin{equation}
  \begin{split} 
    I & =  
    \int\left[f(x, y)-f_{1}(x) f_{2}(y)\right]^{2} d x d y \\
    & = \int f(x, y) d F(x, y)+\int f_{1}(x) d F_{1}(x) \int f_{2}(y) d F_{2}(y) - 2 \int f_{1}(x) f_{2}(y) d F(x, y) \\
    & = E f(X, Y)+E\left[f_{1}(X)\right] E\left[f_{2}(Y)\right]-2 E\left[f_{1}(X) f_{2}(Y)\right] 
  \end{split}
\end{equation}
This lead to the following expression
\begin{equation}
    \hat{I} 
    =\frac{1}{n} \sum_{i=1}^{n} \hat{f}_{-i}\left(X_{i}, Y_{i}\right)
    +\frac{1}{n^{2}} \sum_{i=1}^{n} \sum_{j=1}^{n} \hat{f}_{1,-i}\left(X_{i}\right) \hat{f}_{2,-j}\left(Y_{j}\right) 
    -\frac{2}{n} \sum_{i=1}^{n} \hat{f}_{1,-i}\left(X_{i}\right) \hat{f}_{2,-i}\left(Y_{i}\right) 
\end{equation}
where
\begin{equation}
  \begin{split}
    \hat{f}_{-i}\left(X_{i}, Y_{i}\right) & = \frac{1}{n-1} \sum_{j \neq i} K_{h_{x}}\left(X_{j}-X_{i}\right) K_{h_{y}}\left(Y_{j}-Y_{i}\right) \\ 
    \hat{f}_{1,-i}\left(X_{i}\right) & = \frac{1}{n-1} \sum_{j \neq i} K_{h_{x}}\left(X_{j}-X_{i}\right), \\ 
    \hat{f}_{2,-i}\left(Y_{i}\right) & = \frac{1}{n-1} \sum_{j \neq i} K_{h_{y}}\left(Y_{j}-Y_{i}\right), \\ 
    K_{h_{x}}\left(X_{j}-X_{i}\right) & = \frac{1}{h_{x}} K\left( \frac{X_{i}-X_{j}}{h_{x}} \right), \\ 
    K_{h_{y}}\left(Y_{j}-Y_{i}\right) & = \frac{1}{h_{y}} K\left( \frac{Y_{i}-Y_{j}}{h_{y}} \right).
  \end{split}
\end{equation}

Under some regularity conditions and $H_0$ we have
\begin{equation}
    T  =  \frac{n \left( h_{x} h_{y} \right)^{1/2} \hat{I}}{\hat{\sigma}} \leadsto N(0,1) 
\end{equation}
where
\begin{equation*}
    \hat{\sigma}^{2} = 2\left(n^{2} h_{x} h_{y}\right)^{-1} \sum_{i=1}^{n} \sum_{j \neq i}^{n} K^{2}\left(\frac{X_{i}-X_{j}}{h_{x}}\right) K^{2}\left(\frac{Y_{i}-Y_{j}}{h_{y}}\right) 
\end{equation*}

```{r}
data('faithful')
dat <- data.matrix(faithful)
n <- dim(dat)[1]
h.x <- h.y <- n^(-1/5)


f.hat.remove_i <- function(i){
  idx <- seq(1,n); idx <- idx[-i]
  mean(unlist(lapply(X=idx, FUN=function(jj){dnorm((dat[jj,1]-dat[i,1])/h.x) * dnorm((dat[jj,2]-dat[i,2])/h.y) / h.x / h.y})))
}

f1.hat.remove_i <- function(i){
  idx <- seq(1,n); idx <- idx[-i]
  mean(unlist(lapply(X=idx, FUN=function(jj){dnorm((dat[jj,1]-dat[i,1])/h.x) / h.x })))
}

f2.hat.remove_i <- function(i){
  idx <- seq(1,n); idx <- idx[-i]
  mean(unlist(lapply(X=idx, FUN=function(jj){dnorm((dat[jj,2]-dat[i,2])/h.y) / h.y})))
}

f12.hat.remove_i <- function(i){
  idx <- seq(1,n); idx <- idx[-i]
  mean(unlist(lapply(X=idx, FUN=function(jj){dnorm((dat[jj,1]-dat[i,1])/h.x) / h.x }))) * mean(unlist(lapply(X=idx, FUN=function(jj){dnorm((dat[jj,2]-dat[i,2])/h.y) / h.y})))
}

tmp1 <- mean(unlist(lapply(X=1:n, FUN=f.hat.remove_i)))
tmp2 <- mean(unlist(lapply(X=1:n, FUN=f1.hat.remove_i))) * mean(unlist(lapply(X=1:n, FUN=f2.hat.remove_i)))
tmp3 <- mean(unlist(lapply(X=1:n, FUN=f12.hat.remove_i))) * 2

I <- tmp1 + tmp2 - tmp3
cat('We get I.hat = ', I,'\n')

tmp <- 0
for (ii in 1:n) {
  for (jj in (1:n)[-ii]) {
    tmp <- tmp + (dnorm((dat[ii,1]-dat[jj,1])/h.x))^2 * (dnorm((dat[ii,2]-dat[jj,2])/h.y))^2 
  }
}
sigma.sq <- 2 * tmp / n^2 / h.x / h.y

T <- n * sqrt(h.x*h.y) * I / sqrt(sigma.sq)
cat('Statistic T = ', T,'\n')
```

We got a large statistic $T$ which means we should reject the null hypothesis.

Let's try some different bandwidth selection methods as described in  Homework-10.

```{r}
# bandwith selection
Sig <- cov(dat)
H.plug_in <- diag(diag(sqrt(Sig))) * n^(-1/6) * (4 / 6)^(1/6)
H.scott <- diag(diag(sqrt(Sig))) * n^(-1/6)
```

Firstly, plug-in bandwidth.
```{r}
h.x <- H.plug_in[1,1]; h.y <- H.plug_in[2,2]
tmp1 <- mean(unlist(lapply(X=1:n, FUN=f.hat.remove_i)))
tmp2 <- mean(unlist(lapply(X=1:n, FUN=f1.hat.remove_i))) * mean(unlist(lapply(X=1:n, FUN=f2.hat.remove_i)))
tmp3 <- mean(unlist(lapply(X=1:n, FUN=f12.hat.remove_i))) * 2

I <- tmp1 + tmp2 - tmp3
cat('We get I.hat = ', I,'\n')

tmp <- 0
for (ii in 1:n) {
  for (jj in (1:n)[-ii]) {
    tmp <- tmp + (dnorm((dat[ii,1]-dat[jj,1])/h.x))^2 * (dnorm((dat[ii,2]-dat[jj,2])/h.y))^2 
  }
}
sigma.sq <- 2 * tmp / n^2 / h.x / h.y

T <- n * sqrt(h.x*h.y) * I / sqrt(sigma.sq)
cat('Statistic T = ', T,'\n')
```

Then, Scott bandwidth.
```{r}
h.x <- H.scott[1,1]; h.y <- H.scott[2,2]
tmp1 <- mean(unlist(lapply(X=1:n, FUN=f.hat.remove_i)))
tmp2 <- mean(unlist(lapply(X=1:n, FUN=f1.hat.remove_i))) * mean(unlist(lapply(X=1:n, FUN=f2.hat.remove_i)))
tmp3 <- mean(unlist(lapply(X=1:n, FUN=f12.hat.remove_i))) * 2

I <- tmp1 + tmp2 - tmp3
cat('We get I.hat = ', I,'\n')

tmp <- 0
for (ii in 1:n) {
  for (jj in (1:n)[-ii]) {
    tmp <- tmp + (dnorm((dat[ii,1]-dat[jj,1])/h.x))^2 * (dnorm((dat[ii,2]-dat[jj,2])/h.y))^2 
  }
}
sigma.sq <- 2 * tmp / n^2 / h.x / h.y

T <- n * sqrt(h.x*h.y) * I / sqrt(sigma.sq)
cat('Statistic T = ', T,'\n')
```

In conclusion, we suppose that the *eruptions* and *waiting* are <font face="times" size=5>not independent</font> of each other. 

And also we should notice that different bandwidth selections may greatly influence the statistics $T$. I'd better choice the optimal one.