---
title: "Homework 7"
author: "Liu Huihang"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Problem 1 (Homework 7.1)

Consider Bart Simpson density
\begin{equation}
  f(x)=\frac{1}{2} \phi(x ; 0,1)+\frac{1}{10} \sum_{j=0}^{4} \phi(x ;(j / 2)-1,1 / 10)
\end{equation}
where $\phi(x ;\mu, \sigma)$ denotes a Normal density with mean $\mu$ and standard deviation $\sigma$.

Generate $1000$ sample from Bart Simpson distribution, estimate its density by histogram and compare different interval segmentation method. 
Estimate density use naive density estimator with image and compare the effect of different bandwidth $h$.

## Answer 1

### Data Preparation

We generate samples from multinomial distribution with proportion $0.5, 0.1, 0.1, 0.1, 0.1, 0.1$, whose values correspond to different distributions within the Bart Simpson distribution respectively.

To illustrate the effect of different bandwidth $h$, we can use different breaks in **hist** function. Number of breaks is the number of cells for the histogram which is equivalent to $1/h$.

```{r data}
set.seed(111)
pro <- c(0.5, 0.1, 0.1, 0.1, 0.1, 0.1)
multi <- sample(1:6, 1000, replace = T, prob=pro)
mu <- c(0, seq(-1,1,0.5))
sigma <- c(1, rep(sqrt(1/10), 5))
x <- NULL
for (ii in 1:6) {
  com_txt <- paste("com", ii, " <- rnorm(length(which(multi==", ii, ")), mean=", mu[ii], ", sd=", sigma[ii], ")",sep="")
  eval(parse(text=com_txt))
  com_txt <- paste("x <- c(x, com", ii, ")", sep="")
  eval(parse(text=com_txt))
}
```

### Histogram
```{r}
y <- seq(-4, 4, 0.01)
z <- rep(0, 801)
for (ii in 1:6) {
  z <- z + pro[ii] * dnorm(y, mean=mu[ii], sd=sigma[ii])
}
par(mfrow=c(2,2))
for (breaks in c(10, 30, 70, 120)) {
  hist(x, breaks=breaks, freq=FALSE)
  lines(y, z)
  text(x=2, y=0.4, paste("breaks=", breaks))
}
```

The figure above shows the histogram of samples from Bart Simpson distribution with breaks = $10, 30, 70, 120$, namely, $h=1/10, 1/30, 1/70, 1/120$. 
The solid line in the figure is true density function. 

We find that the histograms underfit the true density function when breaks is small, while overfit the true density function with large breaks. The bar figure fit tight with moderate breaks only. 

### Naive Density Estimator

The naive density estimate is
\begin{equation}
  \hat{f}_h(x) = \frac{1}{hn} \sum_{i=1}^n  K\left( \frac{X_i -x}{h} \right)
\end{equation}
where $K(x)=I(-1<x\le1)/2$.

Let's try $h=0.05, 0.4, 1, 2$ and compare the estimated funciton with true funtion.

```{r naive}
naive <- function(t, h){
  # h is parameter
  # t is variable
  fhx <- 0
  for (xx in x) {
    if (abs(xx - t)/h < 1)
      fhx <- fhx + 1
  }
  fhx / (2 * h * length(x))
}
par(mfrow=c(2,2))
for (h in c(0.05, 0.4, 1, 2)) {
  fh <- unlist(lapply(X=y, FUN=naive, h))
  plot(y, z, "l", lwd=2, col="red", xlab="x", ylab="Density", xlim=c(-4,4), ylim=c(0,0.5))
  lines(y, fh)
  text(x=3, y=0.4, paste("h=",h), )
  title("Naive Density Estimate")
}
```

The red line in the figures is true density function and the blue one is naive density estimator. 
It's the same with the behavior of $h$ in histogram method. Only moderate $h$ fit the true function well.
Too samll or too big $h$ will lead to overfitting or underfitting, respectively.

## Problem 2 (Homework 7.3)
Generate 1000 samples from distribution $0.3N(0,1)+0.7N(1,0.32)$,  draw one figure with different estimated density functions with different bandwidth $h$ where $h$ is given by \emph{kedd}.

## Answer 2

### Data Preparation

We generate samples from binnomial distribution with proportion $0.3, 0.7$, whose values correspond to different distributions within the Bart Simpson distribution respectively.


```{r data2}
library(kedd)
set.seed(111)
pro <- c(0.3, 0.7)
multi <- sample(1:2, 100, replace = T, prob=pro)
mu <- c(0, 1)
sigma <- c(1, 0.3)
x <- NULL
for (ii in 1:2) {
  com_txt <- paste("com", ii, " <- rnorm(length(which(multi==", ii, ")), mean=", mu[ii], ", sd=", sigma[ii], ")",sep="")
  eval(parse(text=com_txt))
  com_txt <- paste("x <- c(x, com", ii, ")", sep="")
  eval(parse(text=com_txt))
}
```

### Bandwidth selection

We conside $h$ to optimize AMISE, maximum likelihood cross-validation,unbiased cross-calidation, biased cross-calidation, complete cross-calidation, modified cross-calidation,trimmed cross-calidation. 

The values of these method is shown following, 


```{r}
amise <- h.amise(x, deriv.order = 0)$h
mlcv <- h.mlcv(x)$h;ucv <- h.ucv(x)$h; bcv <- h.bcv(x)$h; 
ccv <- h.ccv(x)$h; mcv <- h.mcv(x)$h; tcv <-  h.tcv(x)$h
h <- c(amise=amise, mlcv=mlcv, ucv=ucv, bcv=bcv, ccv=ccv, mcv=mcv,tcv=tcv)
print(round(h, 4))
```


### Draw figure

We can generate KDE by **dkde** function with defalut parameters and draw then in one figure as following.

```{r}
y <- seq(-4, 4, 0.01)
z <- rep(0, length(y))
for (ii in 1:2) {
  z <- z + pro[ii] * dnorm(y, mean=mu[ii], sd=sigma[ii])
}
plot(y, z, "l", lty=2, lwd=2, col=1, xlab="x", ylab="Density", xlim=c(-3,4), ylim=c(0,1))
for (ii in 1:7) {
  lines(dkde(x, h=h[ii]), col=ii+1, lwd=2)
}
legend("topright", legend=c("true","amise","mlcv","ucv","bcv","ccv","mcv","tcv"), col=1:8, lty=c(2, rep(1,7)))
title("KDE with different h")
```